---
title: "Stroke Risk Factors"
author: "Kira Wolff"
output: pdf_document
bibliography: bibliography.bib
---
# Introduction

After a stroke occurs, every minute counts [@darehed2020]. The sooner the 
affected person receives medical care, the higher the chances of a good recovery 
and little subsequent damages. Thus it is key to recognize a stroke occured. 
Besides knowledge about the visible symptoms of a stroke, knowledge about who is
at high(er) risk to suffer a stroke can help channeling valuable attention 
resources on the these people.

## Research Question

The goal of this project is to identify medical and/or lifestyle factors that 
are associated with stroke occurence by constructing a model that predicts
stroke occurence and finding the important predictors.

```{r setup, include=FALSE}
# chunk options
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, 
                      message = FALSE, include = FALSE)

# setup seed for reproducibility
set.seed(30051994) 
```

```{r packages}
library(tidyverse) # data handling
library(scales) # add percentages to bar plots
library(ggpubr) # arrange plots
library(caret) # modeling methods
library(randomForest) # to use random forest in caret
library(knitr) # table design
```


# Method

To answer the research question, data about people with and without a stroke in
their medical history will be used. The goal is to build a model that classifies
the people correctly as stroke patients and people without stroke history. This
model could then be applied to other people whose basic demographic,
medical and lifestyle data is known. If the model classifies them as stroke
patients although they did not experience one, they should be treated as high
risk.

Simple comparisons by group will be executed with t-tests or $\chi$²-tests for
continuous and categorial variables, respectively.

As common procedure for binary classification problems, a logistic regression 
will be conducted. 
Considering potentially more complex connections 
between the variables, further modeling approaches will be used. Firstly, a 
regularization term in form of Elastic Net will be added which is especially 
sensible in prediction contexts as it reduces overfitting. Elastic Net was 
choosen instead of LASSO or Ridge regularization because it enables a sort of 
compromise between the other methods, as both parameters can be tuned. 
Secondly, to consider the case that the relationship of the variables is better 
represented by a more complex function, support vector machines (SVMs) with a 
linear kernel will be applied. 
Finally, a classification tree using the random forest algorithm will be 
applied. This might be especially useful to identify the most 
important/impactful variables, that is, predictors. 

To train the models and tune the parameters, repeated 5-fold cross-validation 
will be used. Additionally, the final models' performance will be tested with a 
subset of the data that is not part of the training, consisting of 20% of the 
original dataset. 
The best model will be identified via accuracy as performance criterium. 
Sensitivity and Specifity will also be considered, as well as the general 
complexity of the model - in case of near identical performance, simpler models 
will be chosen.

Finally, the best model will be evaluated regarding the predictors. If possible,
a p-value of 0.05 will be considered as a significant result.

## Implementation

The analysis will be conducted using _R_ version 4.0.3 [@rcoreteam2020] in 
_RStudio_ version 1.3.1093 [@rstudio2020]. Data will be handled via _tidyverse_ 
[@tidyverse2019], primarly with _dplyr_ [dplyr2020], and plotted with _ggplot2_ 
[@ggplot2016], _scales_ [@scales2020] and _ggpubr_ [@ggpubr2020].
The models will be implemented using _caret_ [@caret2020] and _randomForest_ 
[@rforest2002]. For resampling, the methods offered by _caret_ 
will be used. Data will be preprocessed by centering and scaling before building 
the models.


## Data

```{r load and transform data}
# define classes
stroke <- read.csv("healthcare-dataset-stroke-data.csv") %>% 
  mutate(id = as.factor(id),
         gender = as.factor(gender),
         age = as.numeric(age),
         hypertension = as.factor(hypertension),
         heart_disease = as.factor(heart_disease),
         ever_married = as.factor(ever_married),
         work_type = as.factor(work_type),
         Residence_type = as.factor(Residence_type),
         avg_glucose_level = as.numeric(avg_glucose_level),
         bmi = as.numeric(bmi),
         smoking_status = as.factor(smoking_status),
         stroke = as.factor(stroke)
         )
```

```{r impute missing data}
# function to ease identifying variables with missing data
any.na <- function(col){
  any(is.na(col))
}

# identify variables with missing data
apply(stroke[,1:length(names(stroke))], 2, FUN="any.na")


# imputation of missing data via regression
## create dataset without missing data
stroke.complete <- stroke %>% 
  na.omit(bmi)

## build model
lm.bmi <- lm(bmi ~ gender + age + hypertension + heart_disease + ever_married + 
               work_type + Residence_type + avg_glucose_level + 
               smoking_status + stroke,
             data = stroke.complete)

## create dataset with only rows containing missing data
stroke.na <- stroke %>% 
  filter(is.na(bmi))

## predict bmi
bmi.pred <- predict(lm.bmi, stroke.na)

## substitute missing data with predicted data
stroke.na$bmi <- bmi.pred

stroke.clean <- rbind(stroke.na, stroke.complete)
```

```{r check data for irregularities}
summary(stroke.clean)

# Age decimals
sum(stroke.clean$age < 1)
sum(stroke.clean$age < 18)

# bmi extreme
stroke.clean %>% 
  filter(bmi > 50) %>% 
  select(bmi) %>% 
  ggplot(aes(bmi))+
    geom_histogram()

stroke.clean %>% 
  filter(bmi > 70) %>% 
  select(bmi)

# smoking unknown
smok.unk <- round(sum(stroke.clean$smoking_status == "Unknown")/
                    length(stroke.clean$smoking_status), 2)*100

# outcome/stroke ratio
stroke.ratio <- stroke.clean %>%
  summarize(table(stroke)[1]/table(stroke)[2])
```


The data comes from the publicly accessible "Stroke Prediction" dataset which 
was published by the user fedesoriano on kaggle.net in January of 2021 
[@strokedata].
Unfortunately, no further information about the validity of the data is known.
The dataset contains information about demographic, medical and lifestyle data 
from >5000 people.

The independent variable of interest, stroke occurence, is represented in the
data as a binary variable, corresponding to having had a stroke in the own 
medical history or not. Unfortunately, the data is quite unbalanced with a ratio
from `r round(stroke.ratio,2)`:1 for the subjects not affected by a stroke.

Missing Data were handled via single imputation by predicting the missing values
with a linear regression model.

A few variables show disputable values for which more context or explanation by 
the dataset's creator would have been useful. Since this information is not 
provided, I try to evaluate them as appropriate or inappropriate.

Among the age values, multiple cases show values smaller than one. Although this
is not a common way to display age below one year, this is the most obvious
explanation. It is not far-fetched since many other cases are underage as well.
Although the age of people older than one year is just displayed as integer, the
decimals might have been used for infants because an age of zero is even harder 
to interpret or rather incorrect. Additionally, several months make a large 
developmental difference for infants which could be considered here. Lastly, 
reducing the values to "<1" would have removed the possibility to calculate with
age as a continuous variable.

Among the BMI values, outliers reach values up to 97. Although these values are
extreme, they are not impossible and are thus likely valid.

Concerning the smoking status variable, information about `r smok.unk`% of the 
subjects in unknown which makes it difficult to evaluate the contribution of
smoking to the research question. Nevertheless, the other extractable 
information from this variable can be useful.

# Results

This section will describe the sample in more detail, build the models and 
evaluate them.

## Subjects

```{r calculations for descriptives}
nsub <- length(stroke.clean$id)

# demographic
age.min <- min(stroke.clean$age)
age.max <- max(stroke.clean$age)
age.mean <- mean(stroke.clean$age)
age.sd <- sd(stroke.clean$age)

fem.perc <- round(prop.table(table(stroke.clean$gender))[1]*100, 2)
men.perc <- round(prop.table(table(stroke.clean$gender))[2]*100, 2)
div.perc <- round(prop.table(table(stroke.clean$gender))[3]*100, 2)


# health
gluc.mean <- mean(stroke.clean$avg_glucose_level)
gluc.sd <- sd(stroke.clean$avg_glucose_level)

bmi.mean <- mean(stroke.clean$bmi)
bmi.sd <- sd(stroke.clean$bmi)

sick <- stroke.clean %>% 
  filter(hypertension == 1 | heart_disease == 1 | stroke == 1) %>% 
  summarize(n = n(),
            perc = round(n()/nsub*100, 2))


# work
employed <- stroke.clean %>% 
  filter(work_type == "Govt_job" | work_type == "Private" | 
           work_type == "Self-employed") %>%
  summarize(n = n(),
            perc = (n()/nsub)*100)

child <- stroke.clean %>% 
  filter(work_type == "children") %>% 
  summarize(n = n(),
            perc = (n()/nsub)*100)

# smoking 
stroke.smoke <- stroke.clean %>% 
  filter(smoking_status != "Unknown") %>% 
  summarize(prop.table(table(smoking_status))*100)
```

```{r Data Description Plot prep}
# sort variables
var.cat <- c("gender", "hypertension", "heart_disease", "ever_married", 
             "work_type", "Residence_type", "smoking_status", "stroke")
var.cont <- c("age", "avg_glucose_level", "bmi")

# functions to enable "looping" plot creation
plot.bar <- function(data, x){
  ggplot(data, aes_string(x = x))+
    geom_bar(aes(y=..count../sum(..count..)),
             fill = "#4378A2")+
    ylab("Percentage")+
    scale_y_continuous(labels = percent_format())+
    theme_minimal()+
    theme(legend.title = element_blank())
}

plot.hist <- function(data, x){
  ggplot(data, aes_string(x = x))+
    geom_histogram(aes(y=..count../sum(..count..)),
                   fill = "#4378A2")+
    ylab("Percentage")+
    scale_y_continuous(labels = percent_format())+
    theme_minimal()
}

# create base plots
cat.plots <- lapply(var.cat, plot.bar, data = stroke.clean)
cont.plots <- lapply(var.cont, plot.hist, data = stroke.clean)

# refine base plots
plot.gen <- cat.plots[[1]]+
  xlab("Gender")

plot.hyp <- cat.plots[[2]]+
  xlab("Hypertension")+
  scale_x_discrete(labels = c("No", "Yes"))

plot.heart <- cat.plots[[3]]+
  xlab("Heart Disease")+
  scale_x_discrete(labels = c("No", "Yes"))

plot.mar <- cat.plots[[4]]+
  xlab("Married (Lifetime)")

plot.work <- cat.plots[[5]]+
  xlab("Work Type")+
  scale_x_discrete(labels = c("Children", "Govern. \nJob",
                              "Never \nWorked", "Private \nSector", 
                              "Self-\nemployed"))

plot.res <- cat.plots[[6]]+
  xlab("Residence Type")

plot.smok <- cat.plots[[7]]+
  xlab("Smoking Status")+
  scale_x_discrete(labels = c("formerly\nsmoked", "never\nsmoked", 
                              "currently\nsmokes", "unknown"))

plot.stroke <- cat.plots[[8]]+
  xlab("Stroke")+
  scale_x_discrete(labels = c("No", "Yes"))

plot.age <- cont.plots[[1]]+
  xlab("Age")+
  annotate(geom = "text",
           x = 15, y = 0.045, 
           label = paste("M =",
                         round(age.mean, 2),
                         "\nSD =",
                         round(age.sd,2)))


plot.gluc <- cont.plots[[2]]+
  xlab("Glucose Level [mg/dl]")+
  annotate(geom = "text", 
           x = 200, y = 0.1, 
           label = paste("M =",
                         round(gluc.mean, 2),
                         "\nSD =",
                         round(gluc.sd,2)))

plot.bmi <- cont.plots[[3]]+
  xlab("BMI")+
  annotate(geom = "text", 
           x = 75, y = 0.15, 
           label = paste("M =",
                         round(bmi.mean, 2),
                         "\nSD =",
                         round(bmi.sd,2)))

# show prepared plots via separate chunks
```


```{r Plot Demographic, fig.height=2, include=FALSE}
ggarrange(plot.gen, plot.age)
# do not show plot in pdf because text covers information
```

The data represents information about `r nsub` subjects. The sample has a large
age range from <1 to `r age.max` years with a mean of `r round(age.mean, 2)`
years (_SD_ = `r round(age.sd, 2)`). It consists of `r fem.perc`% women,
`r men.perc`% men, and `r div.perc`% identifying with other labels.

```{r Plot Medical, fig.height=6, include=TRUE, fig.cap = "Distributions of Medical Data"}
ggarrange(plot.hyp, plot.heart, plot.stroke, plot.gluc, plot.bmi, ncol=2, 
          nrow=3)
```

As Figure 1 shows, the majority of the sample has no history of hypertension, 
heart disease, stroke or elevated blood glucose level. Excluding the latter, 
`r sick$perc`% are affected by at least one of the aforementioned diagnoses 
overall. 
According to the BMI data, all categories from underweight to obese are 
represented by the sample (though BMI calculations for minors have to be 
interpreted with reservations).

```{r Plot Lifestyle, fig.height=4, include=TRUE, fig.cap = "Distributions of Lifestyle Data"}
ggarrange(plot.mar, plot.work, plot.res, plot.smok)
```

Concerning the lifestyle of the sample, the majority is married or has been at
some point in their life. `r round(employed$perc,2)`% were employed or 
self-employed at the time of data collection, while `r round(child$perc, 2)`% 
were counted as children. 
Regarding their residence, the sample splitted almost fifty-fifty into rural and
urban population. At last, considering only the subsample whose smoking status 
is known, `r round(stroke.smoke[2,],2)`% never smoked, while 
`r round(stroke.smoke[1,],2)`% smoked formerly and 
`r round(stroke.smoke[2,],2)`% currently still do.
More information can be seen in Figure 2.

In general, the sample seems to be quite representative of the general 
population, without focusing on a specific subgroup. Thus, the results of the 
following analysis can potentially be transferred to the general population. 
Although one should keep in mind that some variable connections only appear in 
certain subgroups, so working with samples like this is not always beneficial.

```{r table stroke splitted, include=TRUE }
# calculate descriptives
sample <- stroke.clean %>%
  # split by stroke
  dplyr::group_by(stroke) %>%
  # create summary
  dplyr::summarize(
    # n
    n = n(),
    #gender
    'Gender (m/f)' = paste( 
        round(prop.table(table(gender))[1]*100, 1), 
        "/" ,
        round(prop.table(table(gender))[2]*100, 1), 
        sep=""),
    #age
    Age = paste(
      round(mean(age), 1), 
      " (", 
      round(sd(age), 1), 
      ")", 
      sep=""),
    #hypertension
    'Hypertension (n/y)' = paste( 
        round(prop.table(table(hypertension))[1]*100, 1), 
        "/" ,
        round(prop.table(table(hypertension))[2]*100, 1), 
        sep=""),
    #heart disease
    'Heart Disease (n/y)' = paste( 
        round(prop.table(table(heart_disease))[1]*100, 1), 
        "/" ,
        round(prop.table(table(heart_disease))[2]*100,1), 
        sep=""),
    #glucose
    'Glucose [mg/dl]' = paste(
      round(mean(avg_glucose_level), 1), 
      " (", 
      round(sd(avg_glucose_level), 1), 
      ")", 
      sep=""),
    # bmi
    "BMI" = paste(
      round(mean(bmi), 1),
      " (",
      round(sd(bmi), 1),
      ")",
      sep=""),
    #married
    'Ever Married (n/y)' = paste( 
        round(prop.table(table(ever_married))[1]*100, 1), 
        "/" ,
        round(prop.table(table(ever_married))[2]*100, 1), 
        sep=""),
    #residence
    'Residence (rural/urban)' = paste( 
        round(prop.table(table(Residence_type))[1]*100, 1), 
        "/" ,
        round(prop.table(table(Residence_type))[2]*100, 1), 
        sep=""),
    #smoking
    'Smoking Status (formerly/never/currently)' = paste( 
        round(prop.table(table(smoking_status))[1]*100, 1), 
        "/" ,
        round(prop.table(table(smoking_status))[2]*100, 1),
        "/" ,
        round(prop.table(table(smoking_status))[3]*100, 1), 
        sep=""),
    )

values <- t(sample[-1])

# calculate p-values

sample.sig <- stroke.clean %>% 
  summarize(
    # empty for n
    placeholder = "---",
    # gender
    gender.chi = round(chisq.test(table(gender, stroke))$p.value, 3),
    # age
    age.t = round(t.test(age ~ stroke)$p.value, 3),
    # hypertension
    hyp.chi = round(chisq.test(table(hypertension, stroke))$p.value, 3),
    # heart disease
    heart.chi = round(chisq.test(table(heart_disease, stroke))$p.value, 3),
    # glucose
    gluc.t = round(t.test(avg_glucose_level ~ stroke)$p.value, 3),
    # bmi
    bmi.t = round(t.test(bmi ~ stroke)$p.value, 3),
    # married
    married.chi = round(chisq.test(table(ever_married, stroke))$p.value, 3),
    # residence
    res.chi = round(chisq.test(table(Residence_type, stroke))$p.value, 3),
    # smoking
    smoke.chi = round(chisq.test(table(smoking_status, stroke))$p.value, 3),
  )

p.values <- t(sample.sig)

# change "zero"-p-values to <.001
for(i in 1:length(p.values)){
  if(p.values[i,] == 0){
    p.values[i,] <- "<0.001"
  }
}

table1 <- cbind(values, p.values)

kable(table1,
      col.names= c("No Stroke", "Stroke", "p"),
      caption = "Sample Data by Stroke Occurence (Mean (SD) or Percentages)")
```


A comparison of the variables by stroke, as depicted in Table 1, shows the 
potential of the variables as predictors: For almost all variables, a 
significant difference emerges. Especially age and the variables concerning
health, i.e. hypertension, heart disease, and glucose show large differences.

## Model Evaluation

```{r Data Description}
table(stroke.clean$stroke)

summary(stroke.clean)

sum(stroke.clean$hypertension == 1 | stroke.clean$heart_disease == 1 | 
      stroke.clean$stroke.clean == 1)/length(stroke.clean$id)
```

```{r Model preparation}
# Split Data
# use 80% of the data to fit and tune the models, test accuracy with other 20%
train.index <- sample(1:length(stroke.clean$id),
                      size = floor(length(stroke.clean$id)*0.8))

train.data <- stroke.clean[train.index,]
test.data <- stroke.clean[-train.index,]

# Create Base Model
model <- stroke ~ gender + age + hypertension + heart_disease + ever_married +
  work_type + Residence_type + avg_glucose_level + bmi + smoking_status

# Prepare Cross-validation
trctrl <- trainControl(method = "repeatedcv", 
                       number = 5,
                       repeats = 3)
```



```{r logistic regression}
# fit model
glm.train <- train(model,
                   train.data,
                   trControl = trctrl,
                   method = "glm",
                   preProcess = c("center", "scale"),
                   family = "binomial")

summary(glm.train)
glm.train

# predict test data
glm.preds <- predict(glm.train, test.data)

# compare prediction and real data
glm.mat <- confusionMatrix(glm.preds, test.data$stroke)
```

The logistic regression provides a model with _Acc_ = 
`r round(glm.train$results$Accuracy,3)` which reaches _Acc_ = 
`r round(glm.mat$overall[1],3)` with the test data.

```{r regularized log regression}
# fit model
reg.train <- train(model,
                   train.data,
                   method = "glmnet",
                   preProcess = c("center", "scale"),
                   metric = "Accuracy",
                   tuneLength = 10,
                   family = "binomial")

# save accuracy
reg.train.acc <- reg.train$results %>% 
  filter(alpha == reg.train$bestTune$alpha & 
           lambda == reg.train$bestTune$lambda) %>% 
  summarize(Accuracy = Accuracy)
  

# predict test data
reg.preds <- predict(reg.train, test.data)

# compare prediction and real data
reg.mat <- confusionMatrix(reg.preds, test.data$stroke)

```

The hyperparameters of the logistic regression regularized with Elastic Net were
tuned to $\alpha$ = `r reg.train$bestTune$alpha` and $\lambda$ = 
`r round(reg.train$bestTune$lambda,3)`. 
This provides a model with _Acc_ = 
`r round(reg.train.acc,3)` which reaches _Acc_ = 
`r round(reg.mat$overall[1],3)` with the test data.

```{r SVM}
# fit model
grid <- expand.grid(C = seq(0, 1, 0.05))

svm.train <- train(model,
                   train.data,
                   method = "svmLinear",
                   trControl = trctrl,
                   preProcess = c("center", "scale"),
                   tuneGrid = grid)

# save accuracy
svm.train.acc <- svm.train$results %>% 
  filter(C == svm.train$bestTune$C) %>% 
  summarize(Accuracy = Accuracy)

# predict test data
svm.preds <- predict(svm.train, test.data)

# compare prediction and real data
svm.mat <- confusionMatrix(svm.preds, test.data$stroke)

```

The hyperparameter of the SVM with Elastic Net was
tuned to _C_ = `r svm.train$bestTune$C`.
This provides a model with _Acc_ = 
`r round(svm.train.acc,3)` which reaches _Acc_ = 
`r round(svm.mat$overall[1],3)` with the test data.

```{r Random Forest}
# fit model
tree.train <- train(model,
                   train.data,
                   method = "rf",
                   trControl = trctrl,
                   preProcess = c("center", "scale"),
                   tuneLength = 10)

# save accuracy
tree.train.acc <- tree.train$results %>% 
  filter(mtry == tree.train$bestTune$mtry) %>% 
  summarize(Accuracy = Accuracy)

# predict test data
tree.preds <- predict(tree.train, test.data)

# compare prediction and real data
tree.mat <- confusionMatrix(tree.preds, test.data$stroke)
```

The hyperparameter of the SVM with Elastic Net was
tuned to _mtry_ = `r tree.train$bestTune$mtry`.
This provides a model with _Acc_ = 
`r round(tree.train.acc,3)` which reaches _Acc_ = 
`r round(tree.mat$overall[1],3)` with the test data.

```{r model comparison, include=TRUE}
# list of all models
models <- list(glm.mat, reg.mat, svm.mat, tree.mat)

# function to ease extracting values from list
read.output <- function(list, info){
  if(info == "accuracy"){
    list$overall[1]
  } else { 
    if(info == "sensitivity"){
      list$byClass[1]
      } else {
        if(info == "specifity"){
          list$byClass[2]
        }
      }}}

# dataframe with models' results
model.comp <- data.frame(
  model = c("GLM", "Regularized Regression", "SVM", "Tree"),
  acc = sapply(models, read.output, info="accuracy"),
  sens = sapply(models, read.output, info="sensitivity"),
  spec = sapply(models, read.output, info="specifity")
)

# create nice table
kable(
  model.comp,
  #format = "latex",
  #booktabs = TRUE,
  col.names = c("Model", "Accuracy", "Sensitivity", "Specifity"),
  align = c("l", "c", "c", "c"),
  caption = "Comparison of the Models",
  digits = 3
)
```

The performance of the different models is summarized in Table 2.

## Predictor Evaluation

Since the logistic regression reached highest accuracy while being the simplest 
model, it will be used to take a closer look at the predictors.

```{r GLM predictor}
# save coefficients for text
summary(glm.train)
sum.glm <- summary(glm.train)

age.est <- round(sum.glm$coefficients["age", "Estimate"], 2)
age.p <- round(sum.glm$coefficients["age", "Pr(>|z|)"], 3)

gluc.est <- round(sum.glm$coefficients["avg_glucose_level", "Estimate"], 2)
gluc.p <- round(sum.glm$coefficients["avg_glucose_level", "Pr(>|z|)"], 3)

self.est <- round(sum.glm$coefficients["`work_typeSelf-employed`", 
                                       "Estimate"], 2)
self.p <- round(sum.glm$coefficients["`work_typeSelf-employed`", "Pr(>|z|)"], 3)

urb.est <- round(sum.glm$coefficients["Residence_typeUrban", "Estimate"], 2)
urb.p <- round(sum.glm$coefficients["Residence_typeUrban", "Pr(>|z|)"], 3)
```

The logistic regression with stroke as outcome yields age (_b_ = `r age.est`, 
_p_ < 0.001) and glucose level (_b_ = `r gluc.est`, _p_ = `r gluc.p`) as 
significant predictors. 

This means, that for each year older the odds to suffer a stroke are multiplied
by `r round(exp(age.est),2)`, and for every increase of 1 mg/dl of average blood 
glucose, the odds to suffer a stroke are multiplied by 
`r round(exp(gluc.est),2)`. 
To be precise, these values apply to the "average" person from the sample with 
average values for all variables. The possibility that the odds change 
differently for people with values far from the mean cannot be eliminated with 
this analysis.

Although the other predictors are strictly speaking not significant, the p-value
of two further predictors is close to 0.05. On the one hand, self-employment 
reduces the risk by `r round(exp(self.est)*100,1)`% (_p_ = `r self.p`), on the 
other hand urban residence increases the risk by `r round(exp(urb.est)*100,1)`% 
(_p_ = `r urb.p`).

# Discussion 

According to accuracy as performance criterium, there is no difference between 
logistic regression, regularized logistic regression, SVM and random forest. 
Due to the principle to keep modeling as simple as possible, as long as 
performance does not suffer, logistic regression is to be preferred in
this case. The more complex models did not improve prediction performance. 

On the first glance, the accuracy of `r round(glm.train$results$Accuracy,3)` 
seems to be a good result for the goal to build a model to predict stroke 
occurence. 
Unfortunately, when considering specifity as an additional performance 
criterium, the prediction does not work well. Considering the data, the reason 
for this probably lies in the ratio from subjects with to subjects without 
stroke which is 1:`r round(stroke.ratio,2)`. This ratio leads to an unbalanced 
dataset, and accuracy can be a misleading performance criterium: Because of the 
high portion of healthy subjects, high accuracy is easy to acquire by always 
predicting "no stroke". 
There are solutions for unbalanced datasets like Over- or Under-Sampling which 
unfortunately go beyond the scope of this project work.

Although the created models will thus not work well to predict future stroke
cases, the predictors can still be interpreted. The logistic regression 
presented age as a factor which multiplies the odds to suffer a stroke by 
`r round(exp(age.est),2)` for each passing year. This is not unexpected, as the 
prevalence of many diseases increases with age and stroke is among the most 
common causes for people 65 years of age and older [@sahyoun2001]. 
As second predictor, the average glucose level emerged by multiplying the odds
of a stroke by `r round(exp(gluc.est),2)`. The glucose level is tightly-knit to
the diagnosis of type 2 diabetes which is known to increase stroke risk as well
[@sander2008].

In addition to the significant predictors, the other variables are also
interesting to answer which factors are rather not relevant to assess the 
stroke risk. Here, demographic and medical data seem to be more important than
lifestyle data, though the results for residency type and self-employment are 
interesting. It is possible that some of the other, not-significant
variables correlate with age and thus could not contribute more than age 
already did.


## Limitations

The limitation of this work lies primarly in the unbalanced dataset, which
impeded the model construction with the known methods.

Additionally, the validity of the data is not certain, though no variables or
values stood out in an all too negative way. Also referring to metadata, it 
would have been useful to know how the data where collected, especially
in which city, region, country or culture. With this lack of information,
generalization of the results is limited.

## Future Research

Future research should try to work around the aforementioned limitations, i.e.
use methods to transform the data into a more usable, balanced dataset.
Another possibility would be to use different data that is more balanced to 
begin with.

Although a heterogenous dataset that represents the general population well 
can be important for classification/prediction research, it can also be of use 
to work with a specific subset to identify risk factors of this subset. It would
even be useful to identify which risk factors apply to everyone and which are 
more relevant for specific subgroups.

## Conclusion

Unfortunately, the original goal of this project work to build a well-working
stroke prediction model could not be achieved due to the unbalanced data.
Nevertheless, age and glucose level could be identified as important predictors.


\newpage



# References